---
title: "p8105_hw2_yy3421"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1:

Clean up the pols-month csv:
```{r}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("./data/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

Clean up the snp.csv
```{r}
snp = 
  read_csv("data/snp.csv") |> 
  separate(date, into = c("month", "day", "year"), convert = TRUE) |> 
  arrange(year, month) |> 
  mutate(month = month.name[month]) |> 
  select(year, month, close)
```

Clean up the unemployment.csv:
```{r}
unemployment = 
  read_csv("data/unemployment.csv") |> 
  rename(year = Year) |> 
  pivot_longer(
    Jan:Dec,
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x=_, y = month_df) |> 
  select(year, month, unemployment)
```

Merging the three data frames:

```{r}
data_problem1 = 
  left_join(pols, snp) |> 
  left_join(x = _, y = unemployment)

str(data_problem1)
```

Description:

In the pols-month csv, there are `r nrow(pols)` observations and `r ncol(pols)` variables and tells us about the party affiliation for governors and senators of the time period from `r range(pols$year)[1]` to `r range(pols$year)[2]`. It also presents data of democrat or republican status of the sitting president. The `snp` data has `r nrow(snp)` observations and `r ncol(snp)` variables from years `r range(snp$year)[1]` to `r range(snp$year)[2]`. The `unemployment` data has `r nrow(unemployment)` observations and `r ncol(unemployment)` variables ranging from years `r range(unemployment$year)[1]` to `r range(unemployment$year)[2]`. 

## Problem 2:
```{r}
mr_trash_wheel = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel",
             range = "A2:N586"
             ) |> 
  janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons * 500 / 30,
    trash_wheel_type = "Mr. Trash Wheel",
    year = as.numeric(year)
  )

prof_trash_wheel =
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel",
             range = "A2:M108"
             ) |> 
  janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons * 500 / 30,
    trash_wheel_type = "Professor Trash Wheel"
      )

gwy_trash_wheel = 
    read_excel("data/202309 Trash Wheel Collection Data.xlsx",
             sheet = "Gwynnda Trash Wheel",
             range = "A2:L157"
             ) |> 
  janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons * 500 / 30,
    trash_wheel_type = "Gwynnda Trash Wheel"
      )

```

Combining the three datasets:
```{r}
trash_wheel=
  bind_rows(mr_trash_wheel, prof_trash_wheel, gwy_trash_wheel)
```

Description of the Trash Wheel Dataset:

## Problem 3:

Import and clean baseline dataset:

```{r}
mci_baseline = 
  read_csv("data/data_mci/MCI_baseline.csv",
           skip = 1
           ) |> 
  janitor::clean_names() |> 
  mutate(
    sex = 
      case_match(
        sex,
        1 ~ "male",
        0 ~ "female"),
    sex = as.factor(sex),
    apoe4 =
      case_match(
        apoe4,
        1 ~ "apoe4_carrier",
        0 ~ "apoe4_noncarrier"),
    apoe4 = as.factor(apoe4)
  )
```

Filtering participants who do not meet the stated inclusion criteria:
```{r}
mci_baseline_effective = 
  mutate(
    mci_baseline,
    age_at_onset = 
      ifelse(age_at_onset == ".", NA, age_at_onset)
  ) |> 
  filter(current_age < age_at_onset | is.na(age_at_onset))
```

```{r}
mci_developed =
  mci_baseline_effective |> 
  filter(!(current_age > age_at_onset |is.na(age_at_onset)))
```


Based on the `mci_baseline` dataset, there are **`r nrow(mci_baseline_effective)`** participants recruited for this study. The current_age variable tells us the age of the participant at the study baseline and the age_at_onset variables tells us the participant's age at the onset of MCI. Data on whether the partcipant is an APOE4 carrier or not is also included in the `mci_baseline` dataset. Among these participants, **`r mci_baseline_effective |> filter(!(is.na(age_at_onset))) |>  nrow()`** develop MCI. The average baseline age is `r mci_baseline_effective |> pull(current_age) |> mean() |> round(2)` years old. Among the female participants in this study, `r filter (mci_baseline_effective, sex == "female", apoe4 == "apoe4_carrier") |> nrow()/nrow(filter(mci_baseline_effective, sex == "female"))*100`% are APOE4 carriers.

Import and clean amyloid dataset:
```{r}
mci_amyloid = 
  read_csv("data/data_mci/mci_amyloid.csv",
           skip =1) |> 
  janitor::clean_names() |> 
  rename(
    id = study_id,
  ) |> 
  pivot_longer(
    time_2:time_8,
    names_to = "time_of_data_collected",
    values_to = "time"
  ) 
```

```{r}
mci = 
  inner_join(
    mci_developed, mci_amyloid, by = "id"
  )
```

Based on the two datasets, there are `r mci |> distinct(id) |> nrow()` participants who appear in both datasets. In the `mci` dataset, participants who appear in both datasets are presented with their age, sex, aducaiton, and apoe4 status. Furthermore, each time when the bio-marker is measured is presented under the time variable and the specific time frame which each data was collected is labeled under the time_of_data_collected variable. 
