---
title: "p8105_hw2_yy3421"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1:

Clean up the pols-month csv:
```{r}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("./data/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

Clean up the snp.csv
```{r}
snp = 
  read_csv("data/snp.csv") |> 
  separate(date, into = c("month", "day", "year"), convert = TRUE) |> 
  arrange(year, month) |> 
  mutate(month = month.name[month]) |> 
  select(year, month, close)
```

Clean up the unemployment.csv:
```{r}
unemployment = 
  read_csv("data/unemployment.csv") |> 
  rename(year = Year) |> 
  pivot_longer(
    Jan:Dec,
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x=_, y = month_df) |> 
  select(year, month, unemployment)
```

Merging the three data frames:

```{r}
data_problem1 = 
  left_join(pols, snp) |> 
  left_join(x = _, y = unemployment)

str(data_problem1)
```

Description:

In the pols-month csv, there are `r nrow(pols)` observations and `r ncol(pols)` variables and tells us about the party affiliation for governors and senators of the time period from `r range(pols$year)[1]` to `r range(pols$year)[2]`. It also presents data of democrat or republican status of the sitting president. The `snp` data has `r nrow(snp)` observations and `r ncol(snp)` variables from years `r range(snp$year)[1]` to `r range(snp$year)[2]`. The `unemployment` data has `r nrow(unemployment)` observations and `r ncol(unemployment)` variables ranging from years `r range(unemployment$year)[1]` to `r range(unemployment$year)[2]`. 

## Problem 2:
```{r}
mr_trash_wheel = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel",
             range = "A2:N549"
             ) |> 
  janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons * 500 / 30,
    trash_wheel_type = "Mr. Trash Wheel",
    year = as.numeric(year)
  )

prof_trash_wheel =
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel",
             range = "A2:M108"
             ) |> 
  janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons * 500 / 30,
    trash_wheel_type = "Professor Trash Wheel"
      )

gwy_trash_wheel = 
    read_excel("data/202309 Trash Wheel Collection Data.xlsx",
             sheet = "Gwynnda Trash Wheel",
             range = "A2:L157"
             ) |> 
  janitor::clean_names() |> 
  mutate(
    homes_powered = weight_tons * 500 / 30,
    trash_wheel_type = "Gwynnda Trash Wheel"
      )

```

Combining the three datasets:
```{r}
trash_wheel=
  bind_rows(mr_trash_wheel, prof_trash_wheel, gwy_trash_wheel)
```

